\textbf{Computational Fuzzy Extractors}

### 2013-06-21 CompFuzzExt.tex
changed proof of claim B.5

Let $n$ be a security parameter, $m ,q , \alpha= \poly(n)$.  Let $\beta = \omega(1)$.  All operations are computed $(\mod q)$ and we omit this notation.  
Let $D$ be a distinguisher that breaks $\distLWE_{(m+\alpha), (n+\alpha+\beta), q, \chi'}$ where $\chi'_i = \chi_i$ for $i\leq m$ and $\chi_i \overset{d}= 0$ otherwise.  That is, $D$ can distinguish $(m+\alpha)$ $\LWE$ instances from uniform over $n+\alpha+\beta$ variables where the last $\alpha$ equations have no error added.  We denote by $\vA$ be the uniform distribution over $\Fq^{(m+\alpha)\times(n+\alpha+\beta)}$, $\vx$ be uniform over $\Fq^{(n+\alpha+\beta)}$, and $\ve= (e_1,..., e_{m+\alpha})$ where $e_1,..., e_m$ are generated from error distribution $\chi$ and $e_{m+i} =0$ for $i=1,..,\alpha$. Let $\epsilon \geq 1/\poly(n)$, then, 
\[
|\Pr[D(\vA, \vA\vx+\ve) = 1] - \Pr[D(\vA, \vu) =1]|> \epsilon.
\]
We build a distinguisher that breaks $\distLWE_{m, n, q, \chi}$.  Let $\vA'$ be uniform over $\Fq^{m\times n}$, $\vx'$ be uniform over $\Fq^n$ and $\ve'$ be drawn from $\chi$.  It suffices to build a distinguisher~(of polynomial size) $D'$  where 
\begin{align}
\label{eq:block LWE dist}
|\Pr[D'(\vA', \vA'\vx'+\ve') = 1] - \Pr[D'(\vA', \vu) =1]|> (\epsilon - \ngl(n))(1-\ngl(n)) \approx \epsilon.
\end{align}
We will make a single call to $D$ so we focus on how to prepare a random block fixing instance from the random instance we are given.  The code for $D'$ is given in \figref{fig:perfectLWEreduction}.

\begin{figure}[p]
\begin{framed}
\begin{enumerate}
\item Input $\vA', \vb'$, where $\vA' \overset{\$} \leftarrow \Fq^{m\times n}$ and $\vb'$ is either uniform over $\Fq^m$ or $\vb' = \vA'\vx' +\ve'$.
\item Randomly generate $\vect{R} \overset{\$}\leftarrow \Fq^{\alpha \times n}$. Initialize $\vQ \in \Fq^{n\times (\alpha+\beta)}$ to be the zero matrix.
\item Let $\vb^* = (\vb', b^*_{m+1}, \ldots,b^*_{m+\alpha})$, where $(b^*_{m+1}, \ldots, b^*_{m+\alpha} )\overset{\$} \leftarrow \Fq^\alpha$.\label{step:b generation}
\item Randomly generate $\vect{S} \overset{\$}\leftarrow \Fq^{\alpha \times (\alpha+\beta)}$.
		\subitem If $\rank(\vect{S})<\alpha$, stop and output a random bit.
\item Find a set of $\alpha$ linearly independent columns in $\vS$.  Let $T$ be the set of these indices.
\item For all $x_i$ where $n\leq i$ and $i\not \in T$:\label{step:fill in matrix}
\subitem Generate $x_i\overset{\$}\leftarrow \Fq$.  
\subitem For $j=1,..., m$:
\subsubitem Generate $\vQ_{j, i}\overset{\$}\leftarrow \Fq$
\subsubitem Set $b_j^* = b_j^* + \vQ_{j, i} x_i$.
\item Initialize $\vA^*  = \left(\begin{array}{c | c}\vA' & \vQ\\\hline \vR & \vS\end{array}\right)$.
\item \label{step:randomization}
For {$i=1,..., m$}:
\subitem Randomly generate $\gamma_i \leftarrow \Fq^{1 \times \alpha}$.
\subitem Set $\va_{i} \leftarrow \va^*_{i}+\gamma_i (\vR||\vS)$
\subitem Set $b_i \leftarrow b^*_i + \gamma_i (b^*_{m+1},..., b^*_{m+k})$
\item Run $D$ on input $\vA, \vb$.
\item Output $D$'s output.
\end{enumerate}
\end{framed}
\caption{A PPT $D'$ that distinguishes LWE using distinguisher for LWE w/ block fixing source}
\label{fig:perfectLWEreduction}
\end{figure}
The distinguisher $D'$ has an advantage when $S$ is of rank $\alpha$.  This occurs with overwhelming probability:
\begin{claim}
\label{cl:full rank matrix 5.2}
Let $\vS \overset{\$}\leftarrow \Fq^{\alpha \times (\alpha+\beta)}$ be randomly generated.  Then $\Pr[\rank(\vS)=\alpha]\geq  1- \ngl(n)$.
\end{claim}
\begin{proof}
Direct application of \clref{cl:full rank matrix} when $\alpha = \poly(n)$ and $\beta = \omega(1)$ in $n$.
\end{proof}
\begin{corollary}
$
|\Pr[D(\vA, \vA\vx+\ve) = 1  | \rank(\vS) = \alpha] - \Pr[D(\vA, \vu) =1 | \rank(\vS) = \alpha]|> \epsilon - \ngl(n).
$
\end{corollary}

We are interested in the case where $\rank(\vS) = \alpha$, and it suffices to show that $D'$ prepares a random instance in this case. We show this in the following three claims: 
\begin{enumerate}
\item If $\vA'$ is a random matrix then $\vA$ is a random matrix with $\rank(\vS) = \alpha$.
%\bnote{This should be if $\vb' | A$ is uniformly distributed, then $\vb |A $ is uniformly distributed.}
\item If $\vb' = \vA'\vx'+\ve'$ then $\exists \vx$~(uniformly distributed) where $\vb = \vA \vx + \ve$ and where $\ve_i = \ve_i'$ for $1\leq i\leq m$ and $\ve_i = 0$ otherwise.
\item If $\vb' | \vA'$ is uniformly distributed then $\vb | \vA$ is uniformly distributed.
\end{enumerate}

\begin{claim}
\label{cl:randomMatrixDist}
The matrix $\vA$ generated after Step \ref{step:randomization} is a random matrix with $\rank(\vS) = \alpha$.
\end{claim}
\begin{proof}
It should be clear that $\vR, \vS$ are randomly generated~(conditioned on $\rank(\vS) =\alpha$).  The quadrant represented by $\vA'$ remains random after step~\ref{step:randomization} as we are adding uncorrelated random values.  
The matrix $\vQ_{T^c}$~(recall this is the restriction of $\vQ$ to the columns not in $T$) is random as it is initialized with random values and only uncorrelated random values are added.

Thus, we restrict attention to $\vQ_T$~(the columns of $W$ included in $T$).  Consider a particular row $i$.  Then $\vQ_{i, T} = \gamma_i \vS_{T}$.  Since $\vS_T$ is full rank and $\gamma_i$ is uniformly generated, $\vQ_{i, T}$ is uniformly generated.

Thus, the entire matrix $\vA$ after step~\ref{step:randomization} is uniformly generated conditioned on $\rank(\vS) = \alpha$.

@@

If the pair $\vb' | \vA'$ is uniformly distributed, then $\vb |\vA$ is uniformly distributed.

@@

Let $\vb^*$ be the vector generated after step~\ref{step:fill in matrix}.  The only correlation between $\vb^*$ and $\vA^*$ is in the matrix $\vQ$, i.e. for $1\leq i\leq n$, $b_i^* = b_i + \sum_{n<j \wedge j\not\in T} \vQ_{i, j} x_j$, thus conditioning on $\vQ$, $b_i^*$ is uniform assuming $b_i$ is uniform.  Thus $\vb^* | \vA^*$ is uniformly distributed. We show that $\vb |( \vA, \vA^*)$ is uniformly distributed, therefore, showing that $\vb | \vA$ is uniformly distributed after step~\ref{step:randomization}.  Let $\vgamma$ represent the matrix formed by $\gamma_{i}$.

 The matrix $\vgamma$ is a deterministic function of $(\vA, \vA^*)$.  Thus, $\vb | (\vA, \vA^*) \overset{d}= \vb | (\vA, \vA^*, \vgamma)$.  The only dependence of $\vb$ on $(\vA, \vA^*, \vgamma)$ is in the matrix $\vgamma$.  Thus, $ \vb | (\vA, \vA^*, \vgamma) \overset{d}= \vb | \vgamma$. Recall that 
 \[
 \begin{pmatrix} b_{1}  \\ \vdots \\b_{m} \end{pmatrix}  = \begin{pmatrix} b_{1}^*  \\ \vdots \\b_{m}^* \end{pmatrix}  + \vgamma \begin{pmatrix} b_{m+1}^*  \\ \vdots \\b_{m+\alpha}^* \end{pmatrix}
\]
Thus, $b_1,..., b_m $ are uniformly distributed conditioned on $\vgamma, b_{m+1}^*,..., b_{m+\alpha}^*$.  This completes the claim.
\end{proof}

Finally, the reduction runs in polynomial time and together Claims~\ref{cl:randomMatrixDist},~\ref{cl:random ax+e}, and~\ref{clm:random b} show that when $\rank(\vS) = \alpha$ the distinguisher $D'$ properly prepares the instance thus, 
\begin{align*}
&\left|\Pr\left[D'(\vA', \vu') = 1\right]- \Pr\left[D'(\vA', \vb'=\vA'\vx + \ve)=1\right]  \right|\\
&\, = \left| \Pr\left[D'(\vA', \vu') = 1 | \rank(\vS) = \alpha \right]- \Pr\left[D'(\vA', \vb'=\vA'\vx + \ve)=1 | \rank(\vS) = \alpha\right]\right| \Pr[\rank(\vS) = \alpha] \\
&\, =\left|\Pr\left[D(\vA, \vu)=1 | \rank(S) =\alpha \right]- \Pr\left[D(\vA, \vb)=1 | \rank(S) =\alpha \right]  \right| \Pr[\rank(\vS) = \alpha] \\
&\, \geq (\epsilon - \ngl(n))(1-\ngl(n)) \approx \epsilon
\end{align*}
Thus, Equation~(\ref{eq:block LWE dist}) is satisfied, this completes the proof.
\end{proof}
\section{Additional Proofs}

>>>>

Let $n$ be a security parameter, $m ,q , \alpha= \poly(n)$.  Let $\beta = \omega(1)$.  All operations are computed modulo $q$, and we omit $``\bmod q$'' notation.
Let $D$ be a distinguisher that breaks $\distLWE_{(m+\alpha), (n+\alpha+\beta), q, \chi'}$ with advantage $\epsilon>1/\poly(n)$m
where $\chi'_i = \chi_i$ for $i\leq m$ \lnote{what's $\chi_i$?} and $\chi_i \overset{d}= 0$ otherwise. \lnote{we should probably swap $\chi$ and $\chi'$ to be consistent about the use of primes} Let  $\vA$ denote the uniform distribution over $\Fq^{(m+\alpha)\times(n+\alpha+\beta)}$, $X$ denote the uniform distribution over $\Fq^{(n+\alpha+\beta)}$, and $U$ denote the uniform distribution over $\Fq^{m+\alpha}$ . Then
\[
|\Pr[D(\vA, \vA X+\chi') = 1] - \Pr[D(\vA, U =1]|> \epsilon.
\]
We build a distinguisher that breaks $\distLWE_{m, n, q, \chi}$.  Let $\vA'$ denote the uniform distribution over $\Fq^{m\times n}$, $X'$ denote the uniform distribution over $\Fq^n$, and $U'$ denote the uniform distribution over $\Fq^{m}$ .  We will build a distinguisher $D'$ of polynomial size for which
\begin{align}
\label{eq:block LWE dist}
|\Pr[D'(\vA', \vA'X'+\chi) = 1] - \Pr[D'(\vA', U') =1]|> (\epsilon - \ngl(n))(1-\ngl(n)) \approx \epsilon.
\end{align}
$D'$  will make a single call to $D$, so we focus on how to prepare a random block-fixing instance for $D$ from the random instance that $D'$ is given.  The code for $D'$ is given in \figref{fig:perfectLWEreduction}.

%\begin{figure}[p]
%\begin{framed}
\begin{enumerate}
\item Input $\vA', \vb'$, where $\vA' \overset{\$} \leftarrow \Fq^{m\times n}$ and $\vb'$ is either uniform over $\Fq^m$ or $\vb' = \vA'\vx' +\ve'$ for $\ve'\overset{\$} \leftarrow \chi$ and uniform $\vx'\  \overset{\$} \leftarrow \Fq^n$.
\item Choose $\vect{R} \overset{\$}\leftarrow \Fq^{\alpha \times n}$ uniformly at random. Initialize $\vQ \in \Fq^{m\times (\alpha+\beta)}$  to be the zero matrix.
\item Let $\vb^* = (\vb', b^*_{m+1}, \ldots,b^*_{m+\alpha})$, for uniformly chosen $(b^*_{m+1}, \ldots, b^*_{m+\alpha} )\overset{\$} \leftarrow \Fq^\alpha$.\label{step:b generation}
\item Choose $\vect{S} \overset{\$}\leftarrow \Fq^{\alpha \times (\alpha+\beta)}$ uniformly at random.
		\subitem If $\rank(\vect{S})<\alpha$, stop and output a random bit.
\item Find a set of $\alpha$ linearly independent columns in $\vS$.  Let $T$ be the set of indices of these columns.\label{step:find columns}
\item For all $j\notin T$, $1\le j \le \alpha+\beta$:\lnote{swapped $i$, $j$ and shifted}\label{step:fill in matrix}
\subitem Choose $x_{n+j}\overset{\$}\leftarrow \Fq$ uniformly at random.  
\subitem For $i=1,..., m$:
\subsubitem Choose $\vQ_{i,j}\overset{\$}\leftarrow \Fq$ uniformly at random.
\subsubitem Set $b_i^* = b_i^* + \vQ_{i,j} x_{n+j}$.
\item Initialize $\vA^*  = \left(\begin{array}{c | c}\vA' & \vQ\\\hline \vR & \vS\end{array}\right)$.
\item \label{step:randomization}
For {$i=1,..., m$}:
\subitem Choose a row vector $\gamma_i \leftarrow \Fq^{1 \times \alpha}$ uniformly at random.
\subitem Set $\va_{i} \leftarrow \va^*_{i}+\gamma_i (\vR||\vS)$
\subitem Set $b_i \leftarrow b^*_i + \gamma_i (b^*_{m+1},..., b^*_{m+\alpha})^{\mathsf{T}}$
\item For $i=m+1,\dots, m+\alpha$:
\subitem Set $\va_i \leftarrow \va^*_i$
\subitem Set $b_i = b_i^*$.
\item Output $D(\vA, \vb)$. 
\end{enumerate}
%\end{framed}
%\caption{A PPT $D'$ that distinguishes LWE using distinguisher for LWE w/ block fixing source}
%\label{fig:perfectLWEreduction}
%\end{figure}
The distinguisher $D'$ has an advantage when $\vS$ is of rank $\alpha$.  This occurs with overwhelming probability:
\begin{claim}
\label{cl:full rank matrix 5.2}
Let $\vS \overset{\$}\leftarrow \Fq^{\alpha \times (\alpha+\beta)}$ be randomly generated.  Then $\Pr[\rank(\vS)=\alpha]\geq  1- \ngl(n)$.
\end{claim}
\begin{proof}
Direct result of \clref{cl:full rank matrix} when $\beta = \omega(1)$ in $n$. \lnote{rephrase to say that it follows from the assumption that $q^{-\beta}$ is negligible.}
\end{proof}
\begin{corollary}
$
|\Pr[D(\vA, \vA X+\chi) = 1  | \rank(\vS) = \alpha] - \Pr[D(\vA, U) =1 | \rank(\vS) = \alpha]|> \epsilon - \ngl(n).
$
\end{corollary}
\lnote{why is this a corollary?  A corollary should follow from something above it, usually}

We are interested in the case where $\rank(\vS) = \alpha$, and it suffices to show that $D'$ prepares a good instance for $D$ in this case. We show this in the following three claims: 
\begin{enumerate}
\item If $\vA'$ is a random matrix then $\vA$ is a random matrix subject to the condition that $\rank(\vS) = \alpha$.
%\bnote{This should be if $\vb' | A$ is uniformly distributed, then $\vb |A $ is uniformly distributed.}
\item If $\vb' = \vA'\vx'+\ve'$ for uniform $\vA'$ and $\vx'$, then $\exists \vx$~(uniformly distributed and independent of $\vA$ and $\ve'$) such that $\vb = \vA \vx + \ve$, where $\ve_i = \ve_i'$ for $1\leq i\leq m$ and $\ve_i = 0$ otherwise.
\item If the conditional distribution $\vb'\,|\,\vA'$ is uniform, then the conditional distribution $\vb\,|\,\vA$ is also uniform.
\end{enumerate}

\begin{claim}
\label{cl:randomMatrixDist}
The matrix $\vA$ is distributed as a uniformly random choice from the set of all matrices whose bottom-right $\alpha\times (\alpha+\beta)$ submatrix $\vS$ satisfies $\rank(\vS) = \alpha$.
\end{claim}
\begin{proof}
The bottom $\alpha$ rows of $\vA$ (namely, $\vR|\vS$) are randomly generated~(conditioned on $\rank(\vS) =\alpha$).  The top left $m\times n$ quadrant of $\vA$ is also random, because it is produced as a sum of a uniformly random $\vA'$ with some values that are uncorrelated with $\vA'$.
The submatrix of the top-right $m\times (\alpha+\beta)$ quadrant corresponding to $\vQ_{T^c}$~(recall this is the restriction of $\vQ$ to the columns not in $T$) is also random, because it is initialized with random values to which some uncorrelated values are then added. It is important to note that all these values are independent of $\gamma_i$ values.

Thus, we restrict attention to the $m\times \alpha$ submatrix of $\vA$ that corresponds to $\vQ_T$ in $\vA^*$ (note that these values are $0$ in $\vA^*$).  Consider a particular row $i$. That row is computed as $\gamma_i \vS_{T}$.  Since $\vS_T$ is a full rank square matrix and $\gamma_i$ is uniformly and independently generated, that row is also uniform and independent of other entries in $\vA$.

@@

If the conditional distribution $\vb'\,|\,\vA'$ is uniform, then $\vb\,|\,\vA$ is also uniform.

@@

Since $\vR, \vS$, and $\vQ$ are chosen independently of $\vb'$, the distribution $\vb'\,|\,\vA^*$ is uniform.
Let $\vb^*$ be the vector generated after step~\ref{step:fill in matrix}. Its first $m$ coordinates are  computed by adding the uniform vector $\vb'$ to values that are independent of $\vb^*$, and its remaining $\alpha$ coordinates $b^*_{m+1},\dots,b^*_{m+\alpha}$ are   chosen uniformly.  Thus $\vb^*\,|\,\vA^*$ is uniform.

\lnote{I would delete this paragraph and the next one, and replace with the one below} We will now show that $\vb\,|\,(\vA, \vA^*)$ is uniformly distributed, therefore, showing that $\vb\,|\,\vA$ is uniform.  Indeed, let $\vgamma$ represent the matrix formed by $\gamma_{i}$.

 The matrix $\vgamma$ is a deterministic function of $(\vA, \vA^*)$ \lnote{why?  And does it depend on the choice of the set $T$ in step 5 being deterministic? and do you need this? Can't you just write $\vb\,|\,\vA=\vb\,|\,(\vA^*,\Gamma)$, because $\vA$ is a function of $\vA^*, \Gamma$?}.  Thus, $\vb\,|\,(\vA, \vA^*)$ has the same distribution as $\vb\, |\,(\vA, \vA^*, \vgamma)$.  The only dependence of $\vb$ on $(\vA, \vA^*, \vgamma)$ is in the matrix $\vgamma$\lnote{not sure what the previous sentence means}.  Thus, $ \vb | (\vA, \vA^*, \vgamma) \overset{d}= \vb | \vgamma$. Recall that 
 \[
 \begin{pmatrix} b_{1}  \\ \vdots \\b_{m} \end{pmatrix}  = \begin{pmatrix} b_{1}^*  \\ \vdots \\b_{m}^* \end{pmatrix}  + \vgamma \begin{pmatrix} b_{m+1}^*  \\ \vdots \\b_{m+\alpha}^* \end{pmatrix}
\]
Thus, $b_1,..., b_m $ are uniformly distributed conditioned on $\vgamma, b_{m+1}^*,..., b_{m+\alpha}^*$.  This completes the claim.

\lnote{I got confused by the two paragraphs above --- see comments --- so I wrote it differently.  If you agree, delete the two paragraphs above.}

Let $\vgamma$ represent the matrix formed by $\gamma_{i}$.  It is independent of $\vb^*$ and $\vA^*$, so $\vb^*\,|\,(\vA^*, \vgamma)$ is uniform.    Let $\vgamma'=\left(\begin{array}{c | c}\vect{I_m} & \vgamma \\\hline \vect{0} & \vect{I_\alpha}\end{array}\right)$.
Note that $\vb=\vgamma' \vb^*$.  Since $\vb^*\,|\,(\vA^*, \vgamma)$ is uniform, and $\vgamma'$ is invertible, $\vb\,|\,(\vA^*, \vgamma)$ must also be uniform.
Since $\vA$ is a deterministic function of $\vA^*$ and $\vgamma$ (assuming Step~\ref{step:find columns} is deterministic---if not, we can fix the coins used), the distribution $\vb\,|\,\vA$ is the same  as $\vb\,|\,(\vA^*, \vgamma)$ and is thus also uniform.
\end{proof}

Finally, the reduction runs in polynomial time and together Claims~\ref{cl:randomMatrixDist},~\ref{cl:random ax+e}, and~\ref{clm:random b} show that when $\rank(\vS) = \alpha$ the distinguisher $D'$ properly prepares the instance thus, 
\begin{align*}
&\left|\Pr\left[D'(\vA', \vu') = 1\right]- \Pr\left[D'(\vA', \vb'=\vA'\vx + \ve)=1\right]  \right|\\
&\, = \left| \Pr\left[D'(\vA', \vu') = 1 | \rank(\vS) = \alpha \right]- \Pr\left[D'(\vA', \vb'=\vA'\vx + \ve)=1 | \rank(\vS) = \alpha\right]\right| \Pr[\rank(\vS) = \alpha] \\
&\, =\left|\Pr\left[D(\vA, \vu)=1 | \rank(S) =\alpha \right]- \Pr\left[D(\vA, \vb)=1 | \rank(S) =\alpha \right]  \right| \Pr[\rank(\vS) = \alpha] \\
&\, \geq (\epsilon - \ngl(n))(1-\ngl(n)) \approx \epsilon
\end{align*}
\lnote{change notation above to match notation at the beginning of this section; also, note that the first line follows from teh coinflip if rank is not right} Thus, Equation~(\ref{eq:block LWE dist}) is satisfied, this completes the proof.
\end{proof}
\section{Additional Proofs}
